---
title: "tidymetrics"
author: "Daniel Jiménez"
date: 2020-05-02
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, echo=FALSE, cache=FALSE,warning=FALSE,message=FALSE}
library(tidyverse);
library(tidymetrics);
library(tidymodels);
library(shinybones);
library(shinybones);
library(shinydashboard);
library(scales);
library(lubridate);
library(tidytext);
library(skimr)
theme_set(theme_classic())
```



## Tidymetrics and Shinymetrics

Hi guys! in this opportunity I wanna talk about `tidymetrics`. Why?, Well I have a reason list for say this : **Tidymetrics is a packages that generate the real revolution for Data Science** 
Data science defined as:

>Data science is distinguished from the other [...] fields because its goal is an especially human one: to gain insight and understanding (Robinson,2018)


### Tidymetrics

Tidymetrics is a packages that try to develop new shape for Data analysis through `cross by` functions (periods and dimensions), where aggregated metrics.

### Example with Tidytuesday

```{r, echo=FALSE, cache=FALSE, message=FALSE,warning=FALSE}
tuesdata <- tidytuesdayR::tt_load('2020-04-28')
tuesdata$grosses->grosses
```


### Understanding Data (EDA)


Explore compose data group
```{r}
grosses%>%
  skim()
```



```{r, echo=FALSE, cache=FALSE}
library(tidyverse)
grosses%>%
  group_by(show)%>%
  count(sort = TRUE)%>%
  ungroup()%>%
  top_n(10)%>%
  mutate(show=fct_reorder(show,n))%>%
  ggplot(aes(show,n,fill=factor(show)))+
  geom_col(show.legend = FALSE)+
  labs(x="Show",
       y="Total events")+
  coord_flip()
  
  
  
```


Comportamiento por decada del show


```{r}

## crear el año
grosses%>%
  mutate(year=year(week_ending))%>%
  mutate(decade=(year%/%10)*10)->grosses

grosses%>%
  filter(decade>=1990)%>%
  group_by(decade,show)%>%
  summarize(total_seats=sum(seats_sold),
            capacity=mean(pct_capacity))%>%
  mutate(decade=factor(decade),
         show=reorder_within(show,capacity,decade))%>%
  top_n(20)%>%
  ggplot(aes(show,capacity,fill=decade))+
  geom_col(show.legend = FALSE)+
  facet_wrap(~decade,scales="free_y")+
  scale_x_reordered()+
  scale_y_continuous(expand=c(0,0))+
  labs(y = "Percent seats show occupaded",
         x = NULL,
         title = "Top 15 shows")+
  coord_flip()
```




```{r}
grosses%>%
  filter(show %in% c("Hamilton","The Lion King"))%>%
  ggplot(aes(week_ending,weekly_gross,color=show))+
  geom_line()+
  scale_y_continuous(labels = dollar)+
  expand_limits(y=0)
```


```{r}
the_Lion_summarized<-grosses%>%
  filter(show=="The Lion King")%>%
  rename(date=week_ending)%>%
  cross_by_periods(c("month","quarter","year"))%>%
  summarize(usd_gross=sum(weekly_gross),
            avg_ticket_price=mean(avg_ticket_price))


the_Lion_summarized%>%
  ggplot(aes(date,avg_ticket_price,color=period))+
  geom_line(size=1)
```

