---
title: "tidytuesdat#19"
author: "Daniel Jiménez"
date: 2020-05-11
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I am very excited, becouse is my first time working with Georeferencial Data analysis! In this opportunity I wanna develop only Visual Data Analysis for this project.


Well let's do it!!!

## EDA

I try understand type of events and `eruption_category` in this step.

```{r, echo=FALSE, cache=FALSE, message=FALSE,warning=FALSE}
library(tidyverse);
library(tidymodels);
library(leaflet);
library(maps);
library(patchwork);
library(viridis);
library(tidytext);
library(ggridges);
library(ggthemes);
library(gganimate)
library(anim.plots)

```



The first step to develop is to cross the databases.
```{r, echo=FALSE, cache=FALSE,message=FALSE,warning=FALSE}
## Load Data
volcano <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv')
eruptions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/eruptions.csv')
events <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/events.csv')
```


```{r}
volcano_full<-volcano%>%
  left_join(eruptions,by=c("volcano_number","volcano_name","latitude","longitude"))%>%
  left_join(events,by=c("volcano_number","volcano_name","eruption_number"))

volcano_full$last_eruption_year<-as.numeric(volcano_full$last_eruption_year)
```

Whit this information, I create a visualization per decade of the number of eruptions given the region

```{r, echo=FALSE, cache=FALSE}
volcano_full%>%
  filter(last_eruption_year>=1970 & last_eruption_year!="Unknown")->volcano_full

volcano_full%>%
  mutate(decade = (last_eruption_year%/%10 )*10)->volcano_full

volcano_full%>%
  group_by(region,decade)%>%
  count(sort = TRUE)%>%
  ungroup()%>%
  mutate(decade = as.factor(decade),
           region = reorder_within(region, n, decade)) %>%
  ggplot(aes(region,n,fill=decade))+
  geom_col()+
  facet_wrap(~decade,scales = "free")+
  guides(fill=FALSE)+
  scale_x_reordered() +
  coord_flip()+
  labs(x="Region",
       y="Total Eruptions",
       title = "Eruptions per decade")
  
  
```
 Is incredible this association, maybe is necesary understad (whit this information) the correlation for elevation and number of events
 
 
```{r}
volcano_full%>%
  group_by(region,country)%>%
  summarize(events=n(),
            elevation=mean(elevation,na.rm = TRUE))%>%
  ungroup()%>%
  group_by(region)%>%
  mutate(cor=cor(events,elevation))%>%
  na.omit()%>%
  ggplot(aes(fct_reorder(region,cor),cor,color=cor>0))+
  geom_point(aes(size=cor))+
  coord_flip()+
  labs(x="Region",
       y="Correlation")




```

I think that is possible this relation, maybe is necesary other analysis.


```{r,warning=FALSE,message=FALSE, echo=FALSE, cache=FALSE}
volcano_full%>%
  mutate(year_ago = 2020- start_year)%>%
  ggplot(aes(year_ago +1, fill= event_type))+
  geom_histogram()+
  scale_x_log10()+
  guides(fill=FALSE)
  
```

Well Here We Have the real information

```{r}
volcano_full%>%
  group_by(region,eruption_category)%>%
  count()%>%
  ungroup()%>%
  mutate(eruption_category=factor(eruption_category),
         region=reorder_within(region,n,eruption_category))%>%
  ggplot(aes(region,n))+
  geom_col()+
  scale_x_reordered()+
  facet_wrap(~eruption_category,scales="free",ncol = 1)+
  coord_flip()
    
  
```

Now I make a map

```{r}
volcano_full%>%
  mutate(primary_volcano_type=str_remove(primary_volcano_type,"\\(.*\\)"),
    primary_volcano_type=fct_lump(primary_volcano_type,10))%>%
  ggplot(aes(longitude,latitude))+
  geom_point(aes(color=primary_volcano_type))+
  borders()+
  theme_map()
```

Now with the event type

```{r}
volcano_full%>%
  filter(!is.na(event_type))%>%
  mutate(event_type=str_remove(event_type,"\\(.*\\)"),
         event_type=fct_lump(event_type,5))%>%
  ggplot(aes(longitude,latitude))+
  geom_point(aes(color=event_type),size=.5)+
  borders()+
  theme_map()
```


Well I have one question How are the volcano activity?


```{r,echo=FALSE, cache=FALSE}
volcano_full%>%
  filter(start_year >= 1970)%>%
  ggplot(aes(2020 - start_year))+
  geom_histogram()+
  scale_x_log10()
```

We have more activity in this last years.

and Where?

```{r, echo=FALSE, cache=FALSE}
volcano_full%>%
  mutate(volume=.00001*10^vei)%>%
  ggplot(aes(longitude,latitude))+
  borders()+
  geom_point(aes(size=volume,color=volume))+
  guides(size=FALSE)+
  theme_map()+
  scale_color_gradient2(low="blue",high="red", trans="log10", guide = FALSE)+
  scale_size_continuous(range=c(1,6))

```


And Finally
```{r,warning=FALSE,message=FALSE}
w2hr <- map_data("world") # world map

# Data formatting & analysis ---------

volcanoerupt<-left_join(volcano, eruptions) %>%
  drop_na(c(start_year,vei)) %>%
  filter(start_year>1970, start_year<2019)


plotv<- ggplot() +
  geom_polygon(data=w2hr, 
               aes(x=long, y=lat, group=group), 
               fill="blue", color="red") +
  geom_point(volcano_full, 
             mapping = aes(x = longitude, y = latitude,
                           color = elevation, size=vei,
                           # add a group so the points don't animate from each other
                           group=as.factor(start_year)), 
             shape=17, alpha=0.8)+
  scale_colour_viridis_c("Elevation (m)", option="magma")+
  scale_size_continuous("Volcano \nExplosivity \nIndex")+
  scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0)) +
  coord_fixed(1.3) + theme_minimal() + 
  labs(x = NULL, y = NULL) +
  theme(legend.text = element_text(color="grey30", size=14),
        axis.text = element_blank())+
  theme_map()

plotanimate<-plotv+
  transition_manual(frames=start_year, cumulative = TRUE)+
  ggtitle("Cumulative volcano eruptions per year: {current_frame}")

# render animation
animate(plot = plotanimate, 
  nframes = length(unique(volcano_full$start_year)), 
  fps = 4, end_pause = 8, height = 380, width =600)

## Saving -----------------------------

anim_save("Volcanoes_plot.gif")
```


Thks 4 all Guys!!
